---
title: "CS7200 Spring 25 Logic Grid Insights Project Normal Distribution ANOVA"
author: 'Kaylah Facey and Abdelrahman Madkour'
date: "2025"
output: html_document
---

 
```{r setup, include=FALSE}
#install and load packages here
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(ggplot2)
set.seed(23)
```

#### RUN THIS AND VERIFY EVERYTHING APPEARS IN THE OUTPUT AS EXPECTED BEFORE YOU TURN IT IN

```{r}

# Insights all have a small difference in mu (20s)
constant_delta <- list(
  CROSS_OUT = 30,
  OPENING = 50,
  APPLY_IS = 70,
  APPLY_NOT = 90,
  APPLY_OR = 110,
  APPLY_BEFORE_ONE_SPOT = 130,
  APPLY_BEFORE_N_SPOTS = 150,
  APPLY_BEFORE_UNDEFINED_SPOTS = 170,
  SIMPLE_OR_SAME_CAT = 190,
  SIMPLE_OR_DIFF_CAT = 210,
  BEFORE_DIFF_CAT = 230,
  BEFORE_ONE_SPOT_NOINFO = 250,
  BEFORE_N_SPOTS_NOINFO = 270,
  TRANS_ABC_TRUE = 290,
  TRANS_ABC_FALSE = 310,
  BEFORE_N_SPOTS_SHIFT = 330,
  BEFORE_N_SPOTS_CROSSCHECK = 350,
  TRANS_SETS = 370
)

# Insights are grouped. Within the group the difference in mu is very small (5s) but between groups it is obvious (30s)
grouped_delta <- list(
  # Rules
  CROSS_OUT = 30,
  OPENING = 35,
  # Hints
  APPLY_IS = 65,
  APPLY_NOT = 70,
  APPLY_OR = 75,
  APPLY_BEFORE_ONE_SPOT = 80,
  APPLY_BEFORE_N_SPOTS = 85,
  # Beginner
  APPLY_BEFORE_UNDEFINED_SPOTS = 115,
  SIMPLE_OR_SAME_CAT = 120,
  SIMPLE_OR_DIFF_CAT = 125,
  BEFORE_DIFF_CAT = 130,
  # Intermediate
  BEFORE_ONE_SPOT_NOINFO = 160,
  BEFORE_N_SPOTS_NOINFO = 165,
  TRANS_ABC_TRUE = 170,
  TRANS_ABC_FALSE = 175,
  # Expert
  BEFORE_N_SPOTS_SHIFT = 205,
  BEFORE_N_SPOTS_CROSSCHECK = 210,
  TRANS_SETS = 215
)

muss <- list(constant_delta, grouped_delta)

# Testing with different sds.
sds <- c(10, 30, 60)

# Testing with different numbers of replicates 
rs <- c(2, 5, 10)
```

```{r}
for (mus in muss) {
  for (sd in sds) {
    for (r in rs) {
      simdata<-data.frame(
        a = c(),
        b = c(),
        mu = c()
      );
      for (insight in names(mus)) {
        mu <- mus[insight]
        for (i in 1:r) {
          t <- rnorm(n = 1, mean = as.numeric(mu), sd = as.numeric(sd))
          simdata <- rbind(simdata,c(t, insight))
        }
      }
      colnames(simdata) <- c("Time", "Insight")
      plot <- ggplot(simdata, aes(x=as.numeric(Time), color=Insight, fill = Insight)) + 
        geom_histogram(aes(y=after_stat(density)), position="identity", alpha=0.5, binwidth = 20) +
        geom_density(alpha=0.5) + 
        labs(x = str_interp("Time (mean=${mu}, sd=${sd}, r=${r})"))
      plot
      print(plot)
    }
  }
}
```

#### Additional formatting guidelines: 

For questions where you need to print formulas, please use this formatting: 
$$
y = b_0 + b_1X
$$


how to print a vector:
$$
\begin{bmatrix}
item \\ 
item \\
item
\end{bmatrix}
$$

how to print a matrix:
$$
\begin{bmatrix}
item & item & item\\
item & item & item\\
item & item & item\\
\end{bmatrix}
$$

For questions where you need to print plots, please use clear formatting for title and axes labels (note: you don't have to use ggplot, you can use base R for plotting too).
    
  
Example scatter plot:
```{r}
data(iris)
ggplot(iris) + aes(Petal.Length,Sepal.Length,color= Species) +
  geom_point() + theme_bw() + 
  ggtitle("Relationship between petal and sepal length in iris flower species") + 
  labs(x="Petal length (cm)", y="Sepal length (cm)")

```
