---
title: "CS7200 Spring 25 HW #"
author: '[YOUR NAME]'
date: "01/07/2025"
output: html_document
---


 
```{r setup, include=FALSE}
#install and load packages here
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
```

#### RUN THIS AND VERIFY EVERYTHING APPEARS IN THE OUTPUT AS EXPECTED BEFORE YOU TURN IT IN

#### Problem #1

##### a)
```{r}
#add code here 
insights <- c("CROSS_OUT","OPENING","APPLY_IS","APPLY_NOT","APPLY_OR","APPLY_BEFORE_ONE_SPOT","APPLY_BEFORE_N_SPOTS","APPLY_BEFORE_UNDEFINED_SPOTS","SIMPLE_OR_SAME_CAT","SIMPLE_OR_DIFF_CAT","BEFORE_DIFF_CAT","BEFORE_ONE_SPOT_NOINFO","BEFORE_N_SPOTS_NOINFO","TRANS_ABC_TRUE","TRANS_ABC_FALSE","BEFORE_N_SPOTS_SHIFT","BEFORE_N_SPOTS_CROSSCHECK","TRANS_SETS")
max_time <- 100
df<- data.frame(matrix(nrow=1000,ncol=length(insights)))
colnames(df) <- insights
for (i in 1:length(insights)){
    skewdness <- runif(n=1,min=1,max=7)
    times <- rbeta(n=1000,shape1 = 10, shape2 = skewdness) * max_time 
    df[insights[i]] <- times
}
```

add interpretation and notes here 

##### b)
```{r}
hist(df$CROSS_OUT)
```

##### c)
  
  
#### Additional formatting guidelines: 

  

 wjFor questions where you need to print formulas, please use this formatting: 
$$
y = b_0 + b_1X
$$


how to print a vector:
$$
\begin{bmatrix}
item \\ 
item \\
item
\end{bmatrix}
$$

how to print a matrix:
$$
\begin{bmatrix}
item & item & item\\
item & item & item\\
item & item & item\\
\end{bmatrix}
$$

For questions where you need to print plots, please use clear formatting for title and axes labels (note: you don't have to use ggplot, you can use base R for plotting too).
    
  
Example scatter plot:
```{r}
data(iris)
ggplot(iris) + aes(Petal.Length,Sepal.Length,color= Species) +
  geom_point() + theme_bw() + 
  ggtitle("Relationship between petal and sepal length in iris flower species") + 
  labs(x="Petal length (cm)", y="Sepal length (cm)")

```
